---
title: "DEFT Study 2: from scored data to the published tables"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = "#>")
```

Every Study 2 result in Zweben et al. (2026), in the order the paper runs them,
with the reasoning behind each choice and what changes on cohort 1.

> Zweben, D., Zaneski, L., Chiu, M., Green, H., Brooks-Justice, J., Dachs, A.,
> Vince, M. A., Olino, T., Marshall, P., Steinberg, L., & Chein, J. M. (2026).
> Clarifying the association between digital media engagement and risk taking.
> *Technology in Society, 88*, 103491.
> [doi:10.1016/j.techsoc.2026.103491](https://doi.org/10.1016/j.techsoc.2026.103491)

Study 1 and Study 2 use different measures and different samples. This is Study 2
only: the adolescent and young adult cohorts, PNRT and YRBS risk scores, and the
apple-picking foraging task.

Everything you need to change for a new dataset is in section 1. Nothing after it
refers to a column name from any particular export.

## The order things were run in, and why

The paper builds one argument, and each analysis exists because of what the one
before it left open. Running them out of order still produces numbers, but the
numbers stop answering the question they were meant to answer.

| Paper | Analysis | The question it answers |
|---|---|---|
| 3.2.1 | Descriptives, Table 4 | What do these measures look like in each cohort |
| 3.2.2 | Correlations, Table 5 | Is SSMU related to risk taking and to the traits at all |
| 3.2.3 | Shapley, in text | Of the things that are related, which carries the most weight |
| 3.2.4 | Indirect effects, Table S3 | Do the traits reach SSMU through risk taking |
| 3.2.5 | Subtypes, then hierarchical regression, Table S4 | Which kind of risk, and does any of it survive controlling for traits |
| 3.2.6 | Social vs parasocial, Table 6 | Is this about screen time, or about the kind of platform |

Read down that column and the argument narrows: does risk matter, how much,
through what path, which kind, and on which platforms. Two orderings inside it
are doing real work.

**Shapley comes before the regression, not instead of it.** The predictors are
correlated, so betas depend on entry order. Shapley is order-free, which makes it
the honest first look at relative contribution. The hierarchical regression then
asks the stricter question — whether risk adds anything *after* traits have had
first claim on the shared variance.

**The subtype intercorrelations come before the subtype analyses.** They
establish that the three kinds of risk are related but not interchangeable
(high- and low-stakes negative risk correlate at .55 and .42; low-stakes positive
risk does not track high-stakes negative risk at all). Without that, splitting
risk into three predictors in Step 4 would be unmotivated.

**What changes on cohort 1.** The order does not. What changes is that section 3
has no YRBS substance scoring, so high-stakes negative risk drops out of Table 5,
Table 6, Step 4 of Table S4, and the total risk variety composite. Set
`use_high_stakes <- FALSE` in section 1 and everything downstream adjusts. Step 4
then tests two risk predictors rather than three, which is not comparable to the
published ΔR², so report it on its own terms.

## What you need

```{r packages, eval=FALSE}
install.packages(c("dplyr", "readr", "tidyr", "psych", "Hmisc",
                   "cocor", "mediation", "lm.beta", "ShapleyValue"))
```

```{r libraries}
suppressPackageStartupMessages({
  library(dplyr)
  library(readr)
  library(tidyr)
  library(psych)
  library(Hmisc)
  library(cocor)          # Steiger's z for dependent, overlapping correlations
  library(mediation)      # bootstrapped indirect effects
  library(lm.beta)        # standardized betas
  library(ShapleyValue)   # Shapley value regression
})
```

Put this file next to the data:

```
study-2/
  data/
    study2_analysis_data.csv
    raw/yrbs.csv
    raw/bar.csv
  outputs/                 # created automatically
  deft_study2_pipeline.Rmd
```

A few minutes end to end. Section 9 runs six mediations at 5000 bootstrap
resamples each and takes most of it.

# 1. Configuration

Paths, relative to this file.

```{r paths}
data_dir   <- "data"
output_dir <- "outputs"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

main_file <- "study2_analysis_data.csv"
yrbs_file <- file.path("raw", "yrbs.csv")
bar_file  <- file.path("raw", "bar.csv")
```

The column map. Left side is the name this pipeline uses; right side is the
column in your export. **Change the right side only.** Everything below section 2
uses the names on the left, so this is the only place a column name from a
particular dataset appears.

```{r column-map}
cols <- c(
  pid                = "PID",                                 # also encodes cohort
  gender             = "sex",                                 # 1 female, 2 male
  mtes_time          = "MTES_SocialMediaUse_TimeOnApp.sum",   # sum across platforms
  mtes_checking      = "MTES_PhoneChecking.mean",             # item mean
  mtes_updates       = "MTES_PublicUpdates.mean",             # item mean
  social_use         = "SocialUsage",                         # reciprocal platforms
  parasocial_use     = "ParasocialUsage",                     # one-sided platforms
  impulsivity        = "barrets.sum",                         # BIS, 8 items, sum
  reward_sensitivity = "zuckerman.mean",                      # 6-item subset, mean
  exploration        = "foraging",                            # apples per tree, raw
  risk_low_pos       = "positiverisk_ever_mean",              # PNRT positive, 14 items
  risk_low_neg       = "negativerisk_ever_mean"               # PNRT negative, 7 items
)
```

Cohort split. PID encodes age group — 2xxx adolescents, 3xxx and above young
adults — and that carries over to cohort 1.

```{r cohorts}
is_young_adult <- function(d) d$pid > 3000
is_adolescent  <- function(d) d$pid < 3000
```

High-stakes risk. Seven YRBS substance behaviours collapsed to ever/never;
response coding is not uniform, so they are listed in two groups.

Set `use_high_stakes <- FALSE` if your data has no YRBS substance scoring. Do not
just leave the column missing — total risk variety is built with `na.rm = TRUE`
and would silently average whatever is left.

```{r high-stakes-config}
use_high_stakes <- TRUE

yrbs_never_is_1 <- c("yrbs_alcfirst", "yrbs_mjuse", "yrbs_rxuse",
                     "yrbs_inhalants", "yrbs_cocain")   # 1 = never, anything else = yes
yrbs_yes_is_1   <- c("yrbs_cig_try", "yrbs_vape_use")   # 1 = yes
```

Helpers.

```{r helpers}
zscore <- function(x) (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)

# Blank the CELL, keep the ROW. A participant with an implausible foraging score
# still has valid impulsivity and risk data, so dropping the whole row would
# throw away good observations.
remove_outliers <- function(x, cutoff = 3) {
  z <- zscore(x)
  ifelse(abs(z) > cutoff, NA, x)
}

# Three star levels. Study 1 additionally marked p < .10; Study 2 does not.
format_r <- function(r, p) {
  stars <- if (is.na(p)) "" else if (p < .001) "***" else if (p < .01) "**" else if (p < .05) "*" else ""
  sprintf("%.2f%s", r, stars)
}

format_p_num <- function(p) if (p < .001) "< .001" else sprintf("%.3f", p)

# REDCap exports PIDs as "2002.1" (pid.visit). Harmless if yours are plain.
clean_pid <- function(x) as.numeric(gsub("\\..*$", "", as.character(x)))
```

# 2. Load and rename

Values are untouched. The `setdiff` check fails loudly if a mapped column is not
in the file, which is better than discovering it eight sections later.

```{r load}
raw <- readr::read_csv(file.path(data_dir, main_file), show_col_types = FALSE)

missing_cols <- setdiff(unname(cols), names(raw))
if (length(missing_cols)) {
  stop("Columns in the map not found in ", main_file, ": ", paste(missing_cols, collapse = ", "))
}

d0 <- raw %>%
  dplyr::select(dplyr::all_of(unname(cols))) %>%
  setNames(names(cols)) %>%
  mutate(pid = clean_pid(pid), across(-pid, as.numeric))

nrow(d0)
```

**The SSMU composite is built here, before anything is joined on.** The three MTES
subscales are on different scales — time is a sum across platforms, the other two
are item means — so each is z-scored before averaging. Without that, time would
dominate.

It has to happen at this point rather than later. z-scoring is relative to
whoever is in the frame, so if a join has already changed the row count the mean
and SD shift with it. Section 3 joins the YRBS file, and duplicate PIDs there
will expand the data.

```{r mtes}
d0 <- d0 %>%
  mutate(mtes_zscore = rowMeans(cbind(zscore(mtes_time), zscore(mtes_checking),
                                      zscore(mtes_updates)), na.rm = TRUE))
```

The composite is relative to your sample, so it is not on the same absolute scale
across datasets. Compare patterns of association, not means.

# 3. High-stakes negative risk

Scored from items rather than read in pre-scored. Each behaviour collapses to
ever/never and the score is their mean, so it reads as the proportion of
high-stakes behaviours ever engaged in — the same variety logic as the two PNRT
subtypes, which is what makes the three comparable.

```{r high-stakes}
if (use_high_stakes) {
  yrbs <- readr::read_csv(file.path(data_dir, yrbs_file), show_col_types = FALSE)
  if ("pid_assignment" %in% names(yrbs)) yrbs <- yrbs %>% rename(PID = pid_assignment)
  yrbs <- yrbs %>% mutate(PID = clean_pid(PID))

  # Spelling of the cocaine column varies between exports.
  if (!"yrbs_cocain" %in% names(yrbs) && "yrbs_cocaine" %in% names(yrbs)) {
    yrbs <- yrbs %>% rename(yrbs_cocain = yrbs_cocaine)
  }

  yrbs_missing <- setdiff(c(yrbs_never_is_1, yrbs_yes_is_1), names(yrbs))
  if (length(yrbs_missing)) stop("Missing YRBS items: ", paste(yrbs_missing, collapse = ", "))

  scored <- c(
    lapply(yrbs_never_is_1, function(v) ifelse(!is.na(yrbs[[v]]) & yrbs[[v]] != 1, 1, 0)),
    lapply(yrbs_yes_is_1,   function(v) ifelse(yrbs[[v]] == 1, 1, 0))
  )
  yrbs$risk_high_neg <- rowMeans(as.data.frame(scored), na.rm = TRUE)

  d0 <- d0 %>% left_join(yrbs %>% dplyr::select(PID, risk_high_neg), by = c("pid" = "PID"))
} else {
  d0$risk_high_neg <- NA_real_
}

# The risk subtypes, in the order they appear in the tables.
risk_subtypes <- if (use_high_stakes) {
  c("risk_high_neg", "risk_low_neg", "risk_low_pos")
} else {
  c("risk_low_neg", "risk_low_pos")
}

# Item counts, used to weight the total risk percentage in section 4.
risk_item_counts <- c(risk_high_neg = 7, risk_low_neg = 7, risk_low_pos = 14)

# The tables list the subtypes high-stakes first. Step 4 of the hierarchical
# regression entered them the other way round; order does not change the fit,
# only the row order of the printed table.
risk_entry_order <- rev(risk_subtypes)

risk_subtypes
```

# 4. Descriptives — Table 4

Reported on original scales, so this runs before the composites are built and
before the outlier screen.

Impulsivity needs rescoring here. The models use the BIS sum; Table 4 reports the
mean on the original 1–4 scale (2.88 and 2.87). Four items are worded in the
low-impulsivity direction and are reversed before averaging.

```{r bis-mean}
desc <- d0

bar <- readr::read_csv(file.path(data_dir, bar_file), show_col_types = FALSE)
bar_reverse <- c("bar_dothings", "bar_attention", "bar_saythings", "bar_act")
bar_max <- 4
bar <- bar %>% mutate(across(all_of(bar_reverse), ~ (bar_max + 1) - ., .names = "rev_{col}"))

bar_scored <- c("bar_plantasks", "rev_bar_dothings", "rev_bar_attention",
                "bar_selfcontrol", "bar_concentrate", "bar_careful",
                "rev_bar_saythings", "rev_bar_act")

bar <- bar %>% mutate(impulsivity_mean = rowMeans(across(all_of(bar_scored)), na.rm = TRUE),
                      PID = clean_pid(PID))

desc <- desc %>% left_join(bar %>% dplyr::select(PID, impulsivity_mean), by = c("pid" = "PID"))
```

Total risk variety is weighted by item count, not a plain mean of the three
subtypes. This is why the "Risk Taking, Total Variety" row in Table 4 is not the
average of the rows beneath it.

```{r total-risk-pct}
w <- risk_item_counts[risk_subtypes]
desc$total_risk_pct <- as.numeric(as.matrix(desc[, risk_subtypes]) %*% w / sum(w) * 100)
```

```{r table4}
summarise_var <- function(x) {
  data.frame(Mean = round(mean(x, na.rm = TRUE), 2),
             SD   = round(sd(x,   na.rm = TRUE), 2),
             Min  = round(min(x,  na.rm = TRUE), 2),
             Max  = round(max(x,  na.rm = TRUE), 2))
}

# Row order matches Table 4 top to bottom.
desc_vars <- c("mtes_time", "mtes_updates", "mtes_checking",
               "social_use", "parasocial_use",
               "impulsivity_mean", "reward_sensitivity",
               "total_risk_pct", risk_subtypes, "exploration")

rows <- list()
for (cohort_name in c("Young Adults", "Adolescents")) {
  dfc <- if (cohort_name == "Young Adults") desc[is_young_adult(desc), ] else desc[is_adolescent(desc), ]
  for (v in desc_vars) {
    values <- dfc[[v]]
    # The subtype scores are proportions; Table 4 prints them as percentages.
    if (v %in% risk_subtypes) values <- values * 100
    rows[[length(rows) + 1]] <- data.frame(Cohort = cohort_name, Variable = v,
                                           summarise_var(values), stringsAsFactors = FALSE)
  }
}

table4 <- bind_rows(rows)
write_csv(table4, file.path(output_dir, "study2_table4_descriptives.csv"))
knitr::kable(table4[table4$Cohort == "Young Adults", ], row.names = FALSE)
```

Exploration appears here on its raw apples-per-tree scale, before the sign flip
in section 5. Table 4 describes the task; the models use the flipped version.

The two panel headings in the published Table 4 are transposed — its "Young
Adults" panel holds the adolescent numbers. Age confirms the split here is right
(PID > 3000 is 18–21, PID < 3000 is 11–17), and Table 5 is labelled correctly, so
this is just Table 4.

# 5. Building the remaining analysis variables

**Exploration.** Apples per tree: staying longer on a patch means exploiting, so
higher raw values mean less exploration. Sign-flipped so the variable reads in
the direction of its name. Zeros become NA — a participant who harvested nothing
did not do the task.

```{r exploration}
df <- d0 %>%
  mutate(exploration = ifelse(exploration == 0, NA, exploration),
         exploration = -1 * exploration)
```

**Total risk variety.** Mean of the subtype proportions, then z-scored. The raw
version is kept because section 9 standardizes within its own complete-case
subsample and needs the unstandardized values.

```{r risk-variety}
df <- df %>%
  mutate(risk_variety_raw = rowMeans(dplyr::select(., dplyr::all_of(risk_subtypes)), na.rm = TRUE),
         risk_variety     = zscore(risk_variety_raw))
```

**Outlier screen.** |z| > 3, per variable, cell-wise.

```{r outliers}
analysis_vars <- c("mtes_zscore", "impulsivity", "reward_sensitivity", "exploration",
                   "risk_variety", risk_subtypes, "social_use", "parasocial_use")

df_pre <- df
for (v in analysis_vars) df[[v]] <- remove_outliers(df[[v]])

removed_cells <- sum(sapply(analysis_vars, function(v) sum(!is.na(df_pre[[v]]) & is.na(df[[v]]))))
total_cells   <- sum(sapply(analysis_vars, function(v) sum(!is.na(df_pre[[v]]))))

write_csv(data.frame(removed_cells = removed_cells, total_cells = total_cells,
                     percent_removed = removed_cells / total_cells * 100),
          file.path(output_dir, "study2_outlier_summary.csv"))
write_csv(df, file.path(output_dir, "study2_clean.csv"))
saveRDS(df,   file.path(output_dir, "study2_clean.rds"))

sprintf("%d of %d cells removed (%.3f%%)", removed_cells, total_cells,
        removed_cells / total_cells * 100)
```

# 6. Correlations — Table 5

One published table, two panels, written as two files. Stack them to rebuild
Table 5.

Correlations are pairwise complete, not listwise. Each cell uses everyone who has
both variables, so N differs slightly across cells — which is why the table
carries significance stars rather than a single N.

```{r table5}
main_vars <- c("mtes_zscore", "impulsivity", "reward_sensitivity", "exploration")
row_vars  <- c(main_vars, "risk_variety", risk_subtypes)

get_r_p <- function(d, x, y) {
  d2 <- d[, c(x, y)]
  d2 <- d2[complete.cases(d2), ]
  if (nrow(d2) < 3) return(list(r = NA_real_, p = NA_real_))
  test <- cor.test(d2[[x]], d2[[y]], method = "pearson")
  list(r = unname(test$estimate), p = test$p.value)
}

build_corr_table <- function(d) {
  r_mat <- matrix(NA_real_, nrow = length(row_vars), ncol = length(main_vars),
                  dimnames = list(row_vars, main_vars))
  p_mat <- r_mat
  for (rv in row_vars) for (cv in main_vars) {
    if (rv == cv) next                      # diagonal stays blank, as in the paper
    res <- get_r_p(d, rv, cv)
    r_mat[rv, cv] <- res$r
    p_mat[rv, cv] <- res$p
  }
  formatted <- matrix("", nrow = nrow(r_mat), ncol = ncol(r_mat), dimnames = dimnames(r_mat))
  for (i in seq_len(nrow(r_mat))) for (j in seq_len(ncol(r_mat))) {
    if (!is.na(r_mat[i, j])) formatted[i, j] <- format_r(r_mat[i, j], p_mat[i, j])
  }
  list(formatted = as.data.frame(cbind(Variable = rownames(formatted), formatted)),
       numeric   = as.data.frame(cbind(Variable = rownames(r_mat), r_mat)))
}

res_adults <- build_corr_table(df[is_young_adult(df), ])
write_csv(res_adults$formatted, file.path(output_dir, "study2_table5_correlations_adults.csv"))
write_csv(res_adults$numeric,   file.path(output_dir, "study2_table5_correlations_adults_numeric.csv"))

res_teens <- build_corr_table(df[is_adolescent(df), ])
write_csv(res_teens$formatted, file.path(output_dir, "study2_table5_correlations_teens.csv"))
write_csv(res_teens$numeric,   file.path(output_dir, "study2_table5_correlations_teens_numeric.csv"))

knitr::kable(res_adults$formatted, row.names = FALSE, caption = "Panel A. Young Adults")
```

```{r table5b, echo=FALSE}
knitr::kable(res_teens$formatted, row.names = FALSE, caption = "Panel B. Adolescents")
```

# 7. Risk subtype intercorrelations

Not a table — these open section 3.2.5. They carry the argument that the subtypes
are related but not interchangeable, which is what justifies analysing them
separately in the next section rather than collapsing them.

```{r intercorrelations}
run_corr <- function(d, x, y, cohort) {
  d2 <- d[, c(x, y)]
  d2 <- d2[complete.cases(d2), ]
  if (nrow(d2) < 3) return(data.frame(cohort = cohort, var1 = x, var2 = y, r = NA, p = NA, n = nrow(d2)))
  test <- cor.test(d2[[x]], d2[[y]], method = "pearson")
  data.frame(cohort = cohort, var1 = x, var2 = y,
             r = round(unname(test$estimate), 3),
             p = round(test$p.value, 4), n = nrow(d2))
}

subtype_pairs <- combn(risk_subtypes, 2, simplify = FALSE)

rows <- list()
for (pr in subtype_pairs) rows[[length(rows) + 1]] <- run_corr(df[is_adolescent(df), ],  pr[1], pr[2], "Adolescents")
for (pr in subtype_pairs) rows[[length(rows) + 1]] <- run_corr(df[is_young_adult(df), ], pr[1], pr[2], "Young Adults")

intercor <- bind_rows(rows)
write_csv(intercor, file.path(output_dir, "study2_risk_subtype_intercorrelations.csv"))
knitr::kable(intercor, row.names = FALSE)
```

High- and low-stakes negative risk track each other closely. Low-stakes positive
risk does not track high-stakes negative risk at all.

# 8. Shapley value regression

The predictors are correlated, so ordinary betas depend on entry order and each
predictor's apparent importance shifts with whatever else is in the model.
Shapley averages each predictor's contribution to R² across every possible
ordering, so the result sums to the model R² and does not depend on how the steps
were ordered.

Read it as share of **explained** variance, not effect size. The adolescent model
explains 20.8% of variance in MTES; risk variety accounts for about half of that
20.8%, not half of MTES.

This uses the composite rather than the three subtypes. Three heavily overlapping
risk terms would split one construct's contribution across them and make every
risk share look small. The `na.omit` is listwise by necessity — Shapley refits
over every subset of predictors and needs one constant sample to compare them.

```{r shapley}
run_shapley <- function(data, cohort_label) {
  shap_df <- data %>%
    dplyr::select(mtes_zscore, risk_variety, reward_sensitivity,
                  impulsivity, exploration, gender) %>%
    na.omit()

  shap_values <- shapleyvalue(y = shap_df$mtes_zscore,
                              x = shap_df %>% dplyr::select(-mtes_zscore) %>% as.data.frame())

  # Row 1 = raw Shapley value; row 2 = proportion of model R2, as a percent.
  shap_out <- as.data.frame(shap_values)
  shap_out <- rbind(Shapley_Value = shap_out[1, ],
                    Percent_R2    = round(as.numeric(shap_out[2, ]) * 100, 2))

  write_csv(as.data.frame(shap_out),
            file.path(output_dir, paste0("study2_shapley_", cohort_label, ".csv")))
  shap_out
}

shap_teens  <- run_shapley(df[is_adolescent(df), ],  "adolescents")
shap_adults <- run_shapley(df[is_young_adult(df), ], "adults")
knitr::kable(shap_adults, caption = "Young adults")
```

# 9. Indirect effects — Table S3

Cognitive trait → risk variety → SSMU, separately for each trait and each cohort.
Six mediations, 5000 bootstrap resamples each. Slowest section here.

Risk taking is the mediator and social media engagement is the outcome. The data
are cross-sectional, so this is statistical accounting, not causal sequence. The
paper says "accounted for" throughout and never "led to" — keep that.

Predictors are standardized *after* subsetting to cohort and *after* `na.omit`,
so everything is z-scored within the exact sample the models are fit on.
Standardizing back in section 5 across both cohorts would put adolescents and
young adults on a shared scale and change the coefficients. That is why
`risk_variety_raw` is carried forward.

```{r mediation, results='hide'}
set.seed(2025)   # without this the bootstrap CIs move on every run

run_mediation_set <- function(data, cohort_label) {
  med_df <- data %>%
    dplyr::select(mtes_zscore, risk_variety_raw, reward_sensitivity, impulsivity, exploration) %>%
    mutate(risk_variety_z       = zscore(risk_variety_raw),
           reward_sensitivity_z = zscore(reward_sensitivity),
           impulsivity_z        = zscore(impulsivity),
           exploration_z        = zscore(exploration)) %>%
    na.omit()

  # Two models per mediation: predictor -> mediator, and predictor + mediator ->
  # outcome. Written longhand rather than looped -- mediate() re-evaluates the
  # model calls internally, and building the formulas programmatically breaks it.

  med_model_imp <- lm(risk_variety_z ~ impulsivity_z, data = med_df)
  out_model_imp <- lm(mtes_zscore ~ impulsivity_z + risk_variety_z, data = med_df)
  med_imp <- mediate(med_model_imp, out_model_imp, treat = "impulsivity_z",
                     mediator = "risk_variety_z", boot = TRUE, sims = 5000)

  med_model_rs <- lm(risk_variety_z ~ reward_sensitivity_z, data = med_df)
  out_model_rs <- lm(mtes_zscore ~ reward_sensitivity_z + risk_variety_z, data = med_df)
  med_rs <- mediate(med_model_rs, out_model_rs, treat = "reward_sensitivity_z",
                    mediator = "risk_variety_z", boot = TRUE, sims = 5000)

  med_model_exp <- lm(risk_variety_z ~ exploration_z, data = med_df)
  out_model_exp <- lm(mtes_zscore ~ exploration_z + risk_variety_z, data = med_df)
  med_exp <- mediate(med_model_exp, out_model_exp, treat = "exploration_z",
                     mediator = "risk_variety_z", boot = TRUE, sims = 5000)

  # d0 = indirect (ACME), z0 = direct (ADE), n0 = proportion mediated.
  extract_results <- function(m, predictor_name) {
    data.frame(
      Predictor = predictor_name,
      Indirect  = sprintf("%.2f (%.2f, %.2f), %.3f", m$d0, m$d0.ci[1], m$d0.ci[2], m$d0.p),
      Direct    = sprintf("%.2f (%.2f, %.2f), %.2f", m$z0, m$z0.ci[1], m$z0.ci[2], m$z0.p),
      Prop      = sprintf("%.1f%%, %.2f", m$n0 * 100, m$n0.p),
      stringsAsFactors = FALSE)
  }

  out <- bind_rows(extract_results(med_imp, "impulsivity"),
                   extract_results(med_rs,  "reward_sensitivity"),
                   extract_results(med_exp, "exploration"))

  write_csv(out, file.path(output_dir, paste0("study2_mediation_", cohort_label, ".csv")))
  out
}

med_teens  <- run_mediation_set(df[is_adolescent(df), ],  "adolescents")
med_adults <- run_mediation_set(df[is_young_adult(df), ], "adults")
```

```{r mediation-show, echo=FALSE}
knitr::kable(med_adults, row.names = FALSE, caption = "Young adults")
knitr::kable(med_teens,  row.names = FALSE, caption = "Adolescents")
```

Proportion mediated exceeds 100% for reward sensitivity in both cohorts. That
happens when the direct and indirect paths have opposite signs. It is not an
error and should not be "fixed" — the paper reports it alongside a nonsignificant
p for the proportion.

# 10. Hierarchical regression — Table S4

Once you account for who someone is dispositionally, does what they do still
explain anything about their social media engagement?

1. sex + impulsivity
2. \+ reward sensitivity
3. \+ exploration
4. \+ the risk subtypes

Sex is a covariate from Step 1, not a predictor of interest. Entering the
cognitive traits before risk is the conservative ordering — it gives the
dispositional account first claim on the shared variance, so whatever risk picks
up in Step 4 is over and above traits. The ΔR² at Step 4 is the test.

Listwise deletion happens once, up front, so all four steps are fit on the same
participants. Otherwise the ΔR² comparisons would be across different samples and
`anova()` would be meaningless.

```{r hierarchical}
build_hier_table <- function(data, out_name) {
  model_vars <- c("mtes_zscore", "gender", "impulsivity", "reward_sensitivity",
                  "exploration", risk_entry_order)

  d <- data %>% dplyr::select(dplyr::all_of(model_vars)) %>% na.omit()

  base <- "mtes_zscore ~ gender + impulsivity"
  f <- c(base,
         paste(base, "+ reward_sensitivity"),
         paste(base, "+ reward_sensitivity + exploration"),
         paste(base, "+ reward_sensitivity + exploration +",
               paste(risk_entry_order, collapse = " + ")))

  models     <- lapply(f, function(x) lm(as.formula(x), data = d))
  models_std <- lapply(models, lm.beta)          # betas in the table are standardized
  anova_res  <- do.call(anova, models)           # step-to-step delta R2 tests

  format_beta <- function(beta, p) {
    stars <- if (p < .001) "***" else if (p < .01) "**" else if (p < .05) "*" else if (p < .10) "+" else ""
    sprintf("%.2f%s", beta, stars)
  }

  rows <- list()
  for (i in seq_along(models)) {
    summ <- summary(models[[i]]); summ_std <- summary(models_std[[i]])

    rows[[length(rows) + 1]] <- data.frame(Step = paste0("Step ", i), Predictor = "",
                                           Beta = "", SE = "", t = "", p = "", stringsAsFactors = FALSE)

    for (j in 2:nrow(summ$coefficients)) {       # start at 2 to skip the intercept
      rows[[length(rows) + 1]] <- data.frame(
        Step = "",
        Predictor = rownames(summ$coefficients)[j],
        Beta = format_beta(summ_std$coefficients[j, "Standardized"], summ$coefficients[j, "Pr(>|t|)"]),
        # SE and t come from the unstandardized fit; only beta is standardized.
        SE = sprintf("%.2f", summ$coefficients[j, "Std. Error"]),
        t  = sprintf("%.2f", summ$coefficients[j, "t value"]),
        p  = format_p_num(summ$coefficients[j, "Pr(>|t|)"]),
        stringsAsFactors = FALSE)
    }

    r2 <- summ$r.squared
    fstat <- summ$fstatistic
    fval <- fstat[1]; df1 <- fstat[2]; df2 <- fstat[3]
    pval <- pf(fval, df1, df2, lower.tail = FALSE)

    delta_r2 <- ""
    if (i > 1) {
      delta <- r2 - summary(models[[i - 1]])$r.squared
      delta_r2 <- paste0(", Delta R2 = ", sprintf("%.3f", delta),
                         ", p = ", format_p_num(anova_res$`Pr(>F)`[i]))
    }

    rows[[length(rows) + 1]] <- data.frame(
      Step = "",
      Predictor = paste0("R2 = ", sprintf("%.3f", r2),
                         ", F(", df1, ", ", df2, ") = ", sprintf("%.2f", fval),
                         ", p = ", format_p_num(pval), delta_r2),
      Beta = "", SE = "", t = "", p = "", stringsAsFactors = FALSE)
  }

  out <- bind_rows(rows)
  write_csv(out, file.path(output_dir, out_name))
  out
}

hier_teens  <- build_hier_table(df[is_adolescent(df), ],  "study2_table14_hierarchical_regression_adolescents.csv")
hier_adults <- build_hier_table(df[is_young_adult(df), ], "study2_table13_hierarchical_regression_adults.csv")
knitr::kable(hier_adults, row.names = FALSE, caption = "Young adults")
```

The cohorts diverge here, and it is the substantive finding: for young adults
risk does not add over traits, because impulsivity absorbs it. For adolescents
it does.

**On structuring Step 4.** All subtypes go in together. They correlate between
.25 and .55, so each beta is that subtype's unique contribution net of the
others, and none reaches significance in the young adult model even though the
block as a whole sits near threshold. Two alternatives, both fine if stated:
composite in Step 4 and subtypes in Step 5, which tests "does risk matter" before
"which kind"; or three separate Step 4 models, one subtype each, which reads each
subtype cleanly but runs three tests where the paper ran one. What is not fine is
trying each and reporting whichever makes risk significant.

# 11. Social versus parasocial — Table 6

If the link between social media and risk were just screen time, it would not
matter what kind of platform the time was spent on. It does — risk tracks
reciprocal platforms and not one-sided ones, in both cohorts.

Steiger's z is the right test because the two correlations being compared share a
variable (the risk subtype) and come from the same people. Comparing two r values
directly would ignore that dependency and overstate the difference. The test
needs a third correlation, social use with parasocial use, to know how dependent
they are.

```{r steiger}
run_steiger <- function(data) {
  rows <- list()
  for (pred in risk_subtypes) {
    r_social            <- cor(data[[pred]],    data$social_use,     use = "pairwise.complete.obs")
    r_parasocial        <- cor(data[[pred]],    data$parasocial_use, use = "pairwise.complete.obs")
    r_social_parasocial <- cor(data$social_use, data$parasocial_use, use = "pairwise.complete.obs")
    n <- nrow(na.omit(data[, c(pred, "social_use", "parasocial_use")]))

    test <- cocor.dep.groups.overlap(r.jk = r_social,             # risk x social
                                     r.jh = r_parasocial,         # risk x parasocial
                                     r.kh = r_social_parasocial,  # social x parasocial
                                     n    = n)
    steiger <- test@steiger1980

    rows[[length(rows) + 1]] <- data.frame(
      Risk_Subtype = pred,
      r_social     = round(r_social, 2),
      r_parasocial = round(r_parasocial, 2),
      # Sign flipped so a negative z means social > parasocial, matching the
      # direction printed in Table 6. cocor orders the comparison the other way.
      steiger_z    = round(-1 * steiger$statistic, 2),
      p_value      = round(steiger$p.value, 3),
      n            = n)
  }
  bind_rows(rows)
}

steiger_adults <- run_steiger(df[is_young_adult(df), ])
steiger_teens  <- run_steiger(df[is_adolescent(df), ])
write_csv(steiger_adults, file.path(output_dir, "study2_table7_steiger_adults.csv"))
write_csv(steiger_teens,  file.path(output_dir, "study2_table8_steiger_teens.csv"))

knitr::kable(steiger_adults, row.names = FALSE, caption = "Panel A. Young Adults")
```

```{r steiger-teens, echo=FALSE}
knitr::kable(steiger_teens, row.names = FALSE, caption = "Panel B. Adolescents")
```

# Running this on cohort 1

**High-stakes risk.** Cohort 1 has no YRBS substance scoring. Set
`use_high_stakes <- FALSE` in section 1 and every table adjusts — the row drops
out of Tables 5 and 6, the term drops out of Step 4, and total risk variety
becomes a two-subtype composite. Step 4 then tests two risk predictors rather
than three, so the ΔR² is not comparable to the published one; report it on its
own terms.

Study 1 used the BRP rather than the YRBS for this variable, so if cohort 1
carries a different high-stakes measure there is precedent in the paper for
swapping it in.

**Participants without a phone.** Jason and Lina have already worked out how
people who report no phone at the start of the survey are handled. Confirm with
them that it is taken care of in the cohort 1 export before running anything,
since the MTES composite is z-scored across whoever is in the file.

**Exclusions.** In DEFT we dropped anyone missing one of the at-home cognition
measures — Zuckerman, BIS, foraging. That happened through listwise deletion
inside each model rather than as an explicit filtering step, so N varies a little
between tables.

# Appendix: PNRT ever scores

`risk_low_pos` and `risk_low_neg` arrive already scored. This is how, since the
item-to-valence mapping is not recorded anywhere else.

Every PNRT activity is asked twice: have you ever done this, and if so, how often
in the past six months. The standard variety score uses the second question — an
activity counts as 1 if there was any engagement in the past six months, 0
otherwise.

That was not available. In this version of the survey the frequency follow-up had
no "zero times" option, so once a participant said yes to ever having done
something they could not then report doing it zero times in the past six months.
The frequency item cannot separate recent engagement from none, so variety was
scored from the ever items instead.

If the "zero times" option has since been added, the standard six-month score is
available and is the better measure. It is not the same variable, though — report
it as its own measure rather than as a replication of this one.

Note `pnrt_personal` is scored negative and `pnrt_secret` positive.

```{r pnrt, eval=FALSE}
pnrt_neg_items <- c("pnrt_phone", "pnrt_cheat", "pnrt_skip", "pnrt_personal",
                    "pnrt_snuckout", "pnrt_sexy", "pnrt_mph")            # 7

pnrt_pos_items <- c("pnrt_audition", "pnrt_newclub", "pnrt_truth", "pnrt_food",
                    "pnrt_leader", "pnrt_like", "pnrt_challenge", "pnrt_newlook",
                    "pnrt_newsocial", "pnrt_secret", "pnrt_beliefs",
                    "pnrt_friend", "pnrt_sport", "pnrt_newpeople")       # 14

pnrt <- readr::read_csv(file.path(data_dir, "raw", "pnrt.csv"), show_col_types = FALSE) %>%
  rename(PID = pid_assignment) %>%
  mutate(PID = clean_pid(PID)) %>%
  mutate(negativerisk_ever_mean = rowMeans(across(all_of(pnrt_neg_items)), na.rm = TRUE),
         positiverisk_ever_mean = rowMeans(across(all_of(pnrt_pos_items)), na.rm = TRUE))
```

This mapping reproduces both scored columns exactly on all 254 complete cases.
