---
title: "DEFT Study 2: from scored data to the published tables"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = "#>")
```

Every Study 2 result in Zweben et al. (2026), in the order the paper runs them,
with the reasoning behind each choice and what changes on cohort 1.

> Zweben, D., Zaneski, L., Chiu, M., Green, H., Brooks-Justice, J., Dachs, A.,
> Vince, M. A., Olino, T., Marshall, P., Steinberg, L., & Chein, J. M. (2026).
> Clarifying the association between digital media engagement and risk taking.
> *Technology in Society, 88*, 103491.
> [doi:10.1016/j.techsoc.2026.103491](https://doi.org/10.1016/j.techsoc.2026.103491)

Study 1 and Study 2 use different measures and different samples. This is Study 2
only: the adolescent and young adult cohorts, PNRT and YRBS risk scores, and the
apple-picking foraging task.

Everything you need to change for a new dataset is in section 1. Nothing after it
refers to a column name from any particular export.

## The order things were run in, and why

The paper builds one argument, and each analysis exists because of what the one
before it left open. Running them out of order still produces numbers, but the
numbers stop answering the question they were meant to answer.

| Paper | Analysis | The question it answers |
|---|---|---|
| 3.2.1 | Descriptives, Table 4 | What do these measures look like in each cohort |
| 3.2.2 | Correlations, Table 5 | Is SSMU related to risk taking and to the traits at all |
| 3.2.3 | Shapley, in text | Of the things that are related, which carries the most weight |
| 3.2.4 | Indirect effects, Table S3 | Do the traits reach SSMU through risk taking |
| 3.2.5 | Subtypes, then hierarchical regression, Table S4 | Which kind of risk, and does any of it survive controlling for traits |
| 3.2.6 | Social vs parasocial, Table 6 | Is this about screen time, or about the kind of platform |

Read down that column and the argument narrows: does risk matter, how much,
through what path, which kind, and on which platforms. Two orderings inside it
are doing real work.

**Shapley comes before the regression, not instead of it.** The predictors are
correlated, so betas depend on entry order. Shapley is order-free, which makes it
the honest first look at relative contribution. The hierarchical regression then
asks the stricter question — whether risk adds anything *after* traits have had
first claim on the shared variance.

**The subtype intercorrelations come before the subtype analyses.** They
establish that the three kinds of risk are related but not interchangeable
(high- and low-stakes negative risk correlate at .55 and .42; low-stakes positive
risk does not track high-stakes negative risk at all). Without that, splitting
risk into three predictors in Step 4 would be unmotivated.

**What changes on cohort 1.** The order does not. What changes is that there is
no YRBS substance scoring, so high-stakes negative risk drops out of Table 5,
Table 6, Step 4 of Table S4, and the total risk variety composite. Set
`use_high_stakes <- FALSE` in section 1 and everything downstream adjusts. Step 4
then tests two risk predictors rather than three, which is not comparable to the
published ΔR², so report it on its own terms.

## Scoring and cleaning are yours

This starts from an analysis-ready file: one row per participant, every scale
already scored, outliers already handled. It does not score anything.

DEFT used a good many more scales than show up in these models, and they were all
scored to the lab's standards. Rebuilding a couple of them inside an analysis
script would only ever be a partial picture of that, so it is left out entirely.
Score it the way the lab scores it.

Same for outlier handling. For reference, DEFT screened at |z| > 3 per variable
and blanked the offending *cell* rather than dropping the row, so a participant
with one implausible foraging value still contributed everywhere else. That
removed about half a percent of cells. Use whatever your standard is.

If any of the scoring is unclear, ask me — happy to sit down and go through it
rather than have it guessed at.

## PNRT: important caveat

Scoring is yours, but the PNRT is worth writing down because we scored it
differently than usual and the item lists are not recorded anywhere else.

Every PNRT activity is asked twice: have you ever done this, and if so, how often
in the past six months. Variety is normally scored off the second question — an
activity counts as 1 if there was any engagement in the past six months and 0 if
there was none.

We could not do that. For the first six months of LITE — before Avi and I were on
it — the frequency follow-up gave participants no way to answer zero times in the
past six months once they had already said yes to ever having done the activity.
So for those participants the frequency item cannot tell recent engagement apart
from none.

We score variety off the ever items, and we have kept doing that since. The zero
option may well have been fixed by now, but do not switch — the ever score is
what DEFT uses, and cohort 1 needs to be scored the same way for the two to be
comparable.

`pnrt_personal` counts as negative and `pnrt_secret` counts as positive, which is
not what the names suggest.

```{r pnrt, eval=FALSE}
# Not run -- reference only.
pnrt_neg_items <- c("pnrt_phone", "pnrt_cheat", "pnrt_skip", "pnrt_personal",
                    "pnrt_snuckout", "pnrt_sexy", "pnrt_mph")            # 7

pnrt_pos_items <- c("pnrt_audition", "pnrt_newclub", "pnrt_truth", "pnrt_food",
                    "pnrt_leader", "pnrt_like", "pnrt_challenge", "pnrt_newlook",
                    "pnrt_newsocial", "pnrt_secret", "pnrt_beliefs",
                    "pnrt_friend", "pnrt_sport", "pnrt_newpeople")       # 14

pnrt <- readr::read_csv(file.path(data_dir, "raw", "pnrt.csv"), show_col_types = FALSE) %>%
  rename(PID = pid_assignment) %>%
  mutate(PID = clean_pid(PID)) %>%
  mutate(negativerisk_ever_mean = rowMeans(across(all_of(pnrt_neg_items)), na.rm = TRUE),
         positiverisk_ever_mean = rowMeans(across(all_of(pnrt_pos_items)), na.rm = TRUE))
```

This mapping reproduces both scored columns exactly on all 254 complete cases.

## What you need

```{r packages, eval=FALSE}
install.packages(c("dplyr", "readr", "tidyr", "psych", "Hmisc",
                   "cocor", "mediation", "lm.beta", "ShapleyValue"))
```

```{r libraries}
suppressPackageStartupMessages({
  library(dplyr)
  library(readr)
  library(tidyr)
  library(psych)
  library(Hmisc)
  library(cocor)          # Steiger's z for dependent, overlapping correlations
  library(mediation)      # bootstrapped indirect effects
  library(lm.beta)        # standardized betas
  library(ShapleyValue)   # Shapley value regression
})
```

Put this file next to the data:

```
study-2/
  data/study2_scored.csv     # your scored, cleaned, one-row-per-participant file
  outputs/                   # created automatically
  deft_study2_pipeline.Rmd
```

A few minutes end to end. Section 9 runs six mediations at 5000 bootstrap
resamples each and takes most of it.

# 1. Configuration

Paths, relative to this file.

```{r paths}
data_dir   <- "data"
output_dir <- "outputs"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

main_file <- "study2_scored.csv"
```

The column map. Left side is the name this pipeline uses; right side is the
column in your export. **Change the right side only.** Everything below section 2
uses the names on the left, so this is the only place a column name from a
particular dataset appears.

```{r column-map}
cols <- c(
  pid                = "PID",                                 # also encodes cohort
  gender             = "sex",                                 # 1 female, 2 male
  mtes_time          = "MTES_SocialMediaUse_TimeOnApp.sum",   # sum across platforms
  mtes_checking      = "MTES_PhoneChecking.mean",             # item mean
  mtes_updates       = "MTES_PublicUpdates.mean",             # item mean
  social_use         = "SocialUsage",                         # reciprocal platforms
  parasocial_use     = "ParasocialUsage",                     # one-sided platforms
  impulsivity        = "barrets.sum",                         # BIS, 8 items, sum
  reward_sensitivity = "zuckerman.mean",                      # 6-item subset, mean
  exploration        = "foraging",                            # apples per tree, raw
  risk_low_pos       = "positiverisk_ever_mean",              # PNRT positive, 14 items
  risk_low_neg       = "negativerisk_ever_mean",              # PNRT negative, 7 items
  risk_high_neg      = "yrbs_var"                             # YRBS substance, 7 items
)
```

Exploration is the only one of these that is not a finished score — it comes in
as the raw task output and gets sign-flipped in section 4.

**The platform split.** `social_use` and `parasocial_use` are time on two groups
of platforms, and the whole of Table 6 rests on the distinction:

| | Platforms |
|---|---|
| **Social** — reciprocal, you interact with people you know | Instagram, Snapchat, Facebook |
| **Parasocial** — one-sided, you consume what strangers made | YouTube, TikTok, Pinterest, Twitter |

Cohort 1 will have platforms these lists do not cover. Sort any new one by
whether the typical use is reciprocal interaction or one-way consumption, and
write down what you decided — that call is a finding, not a formatting detail.

Cohort split. PID encodes age group — 2xxx adolescents, 3xxx and above young
adults — and that carries over to cohort 1.

```{r cohorts}
is_young_adult <- function(d) d$pid > 3000
is_adolescent  <- function(d) d$pid < 3000
```

Set `use_high_stakes <- FALSE` if you have no high-stakes risk score. Do not just
leave the column out of the map — total risk variety is built with
`na.rm = TRUE`, so a missing column would not error, it would quietly average
whatever is left.

```{r high-stakes-config}
use_high_stakes <- TRUE

if (!use_high_stakes) cols <- cols[names(cols) != "risk_high_neg"]
```

Helpers.

```{r helpers}
zscore <- function(x) (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)

# Three star levels. Study 1 additionally marked p < .10; Study 2 does not.
format_r <- function(r, p) {
  stars <- if (is.na(p)) "" else if (p < .001) "***" else if (p < .01) "**" else if (p < .05) "*" else ""
  sprintf("%.2f%s", r, stars)
}

format_p_num <- function(p) if (p < .001) "< .001" else sprintf("%.3f", p)

# REDCap exports PIDs as "2002.1" (pid.visit). Harmless if yours are plain.
clean_pid <- function(x) as.numeric(gsub("\\..*$", "", as.character(x)))
```

# 2. Load and rename

Values are untouched. The `setdiff` check fails loudly if a mapped column is not
in the file, which is better than discovering it eight sections later.

```{r load}
raw <- readr::read_csv(file.path(data_dir, main_file), show_col_types = FALSE)

missing_cols <- setdiff(unname(cols), names(raw))
if (length(missing_cols)) {
  stop("Columns in the map not found in ", main_file, ": ", paste(missing_cols, collapse = ", "))
}

d0 <- raw %>%
  dplyr::select(dplyr::all_of(unname(cols))) %>%
  setNames(names(cols)) %>%
  mutate(pid = clean_pid(pid), across(-pid, as.numeric))

nrow(d0)
```

**The SSMU composite is built here, before anything is joined on.** The three MTES
subscales are on different scales — time is a sum across platforms, the other two
are item means — so each is z-scored before averaging. Without that, time would
dominate.

It has to happen at this point rather than later. z-scoring is relative to
whoever is in the frame, so if a join has already changed the row count the mean
and SD shift with it. Section 3 joins the YRBS file, and duplicate PIDs there
will expand the data.

```{r mtes}
d0 <- d0 %>%
  mutate(mtes_zscore = rowMeans(cbind(zscore(mtes_time), zscore(mtes_checking),
                                      zscore(mtes_updates)), na.rm = TRUE))
```

The composite is relative to your sample, so it is not on the same absolute scale
across datasets. Compare patterns of association, not means.

# 3. Risk subtypes

```{r subtypes}
if (!use_high_stakes) d0$risk_high_neg <- NA_real_

# The risk subtypes, in the order they appear in the tables.
risk_subtypes <- if (use_high_stakes) {
  c("risk_high_neg", "risk_low_neg", "risk_low_pos")
} else {
  c("risk_low_neg", "risk_low_pos")
}

# Item counts, used to weight the total risk percentage in section 4. Each
# subtype score is a proportion over a different number of items.
risk_item_counts <- c(risk_high_neg = 7, risk_low_neg = 7, risk_low_pos = 14)

# The tables list the subtypes high-stakes first. Step 4 of the hierarchical
# regression entered them the other way round; order does not change the fit,
# only the row order of the printed table.
risk_entry_order <- rev(risk_subtypes)

risk_subtypes
```

# 4. Descriptives — Table 4

Reported on original scales, so this runs before the composites are built.

```{r desc-setup}
desc <- d0

# Total risk variety, as a percentage. Weighted by item count rather than a plain
# mean of the subtypes, which is why the "Risk Taking, Total Variety" row in
# Table 4 is not the average of the rows beneath it.
w <- risk_item_counts[risk_subtypes]
desc$total_risk_pct <- as.numeric(as.matrix(desc[, risk_subtypes]) %*% w / sum(w) * 100)
```

The published Table 4 reports impulsivity as a mean on the BIS 1–4 scale (2.88
and 2.87) while the models use the sum. This reports whatever you supplied as
`impulsivity`, so say in your table note which one it is.

```{r table4}
summarise_var <- function(x) {
  data.frame(Mean = round(mean(x, na.rm = TRUE), 2),
             SD   = round(sd(x,   na.rm = TRUE), 2),
             Min  = round(min(x,  na.rm = TRUE), 2),
             Max  = round(max(x,  na.rm = TRUE), 2))
}

# Row order matches Table 4 top to bottom.
desc_vars <- c("mtes_time", "mtes_updates", "mtes_checking",
               "social_use", "parasocial_use",
               "impulsivity", "reward_sensitivity",
               "total_risk_pct", risk_subtypes, "exploration")

rows <- list()
for (cohort_name in c("Young Adults", "Adolescents")) {
  dfc <- if (cohort_name == "Young Adults") desc[is_young_adult(desc), ] else desc[is_adolescent(desc), ]
  for (v in desc_vars) {
    values <- dfc[[v]]
    # The subtype scores are proportions; Table 4 prints them as percentages.
    if (v %in% risk_subtypes) values <- values * 100
    rows[[length(rows) + 1]] <- data.frame(Cohort = cohort_name, Variable = v,
                                           summarise_var(values), stringsAsFactors = FALSE)
  }
}

table4 <- bind_rows(rows)
write_csv(table4, file.path(output_dir, "study2_table4_descriptives.csv"))
knitr::kable(table4[table4$Cohort == "Young Adults", ], row.names = FALSE)
```

Exploration appears here on its raw apples-per-tree scale, before the sign flip
in section 5. Table 4 describes the task; the models use the flipped version.

The two panel headings in the published Table 4 are transposed — its "Young
Adults" panel holds the adolescent numbers. Age confirms the split here is right
(PID > 3000 is 18–21, PID < 3000 is 11–17), and Table 5 is labelled correctly, so
this is just Table 4.

# 5. Building the remaining analysis variables

**Exploration.** Apples per tree: staying longer on a patch means exploiting, so
higher raw values mean *less* exploration. Sign-flipped so the variable reads in
the direction of its name. Zeros become NA — a participant who harvested nothing
did not do the task.

```{r exploration}
df <- d0 %>%
  mutate(exploration = ifelse(exploration == 0, NA, exploration),
         exploration = -1 * exploration)
```

**Total risk variety.** Mean of the subtype proportions, then z-scored. The raw
version is kept because section 9 standardizes within its own complete-case
subsample and needs the unstandardized values.

```{r risk-variety}
df <- df %>%
  mutate(risk_variety_raw = rowMeans(dplyr::select(., dplyr::all_of(risk_subtypes)), na.rm = TRUE),
         risk_variety     = zscore(risk_variety_raw))

write_csv(df, file.path(output_dir, "study2_analysis_variables.csv"))
nrow(df)
```

That is everything derived. No outlier screening happens here — see above.

# 6. Correlations — Table 5

One published table, two panels, written as two files. Stack them to rebuild
Table 5.

Correlations are pairwise complete, not listwise. Each cell uses everyone who has
both variables, so N differs slightly across cells — which is why the table
carries significance stars rather than a single N.

```{r table5}
main_vars <- c("mtes_zscore", "impulsivity", "reward_sensitivity", "exploration")
row_vars  <- c(main_vars, "risk_variety", risk_subtypes)

get_r_p <- function(d, x, y) {
  d2 <- d[, c(x, y)]
  d2 <- d2[complete.cases(d2), ]
  if (nrow(d2) < 3) return(list(r = NA_real_, p = NA_real_))
  test <- cor.test(d2[[x]], d2[[y]], method = "pearson")
  list(r = unname(test$estimate), p = test$p.value)
}

build_corr_table <- function(d) {
  r_mat <- matrix(NA_real_, nrow = length(row_vars), ncol = length(main_vars),
                  dimnames = list(row_vars, main_vars))
  p_mat <- r_mat
  for (rv in row_vars) for (cv in main_vars) {
    if (rv == cv) next                      # diagonal stays blank, as in the paper
    res <- get_r_p(d, rv, cv)
    r_mat[rv, cv] <- res$r
    p_mat[rv, cv] <- res$p
  }
  formatted <- matrix("", nrow = nrow(r_mat), ncol = ncol(r_mat), dimnames = dimnames(r_mat))
  for (i in seq_len(nrow(r_mat))) for (j in seq_len(ncol(r_mat))) {
    if (!is.na(r_mat[i, j])) formatted[i, j] <- format_r(r_mat[i, j], p_mat[i, j])
  }
  list(formatted = as.data.frame(cbind(Variable = rownames(formatted), formatted)),
       numeric   = as.data.frame(cbind(Variable = rownames(r_mat), r_mat)))
}

res_adults <- build_corr_table(df[is_young_adult(df), ])
write_csv(res_adults$formatted, file.path(output_dir, "study2_table5_correlations_adults.csv"))
write_csv(res_adults$numeric,   file.path(output_dir, "study2_table5_correlations_adults_numeric.csv"))

res_teens <- build_corr_table(df[is_adolescent(df), ])
write_csv(res_teens$formatted, file.path(output_dir, "study2_table5_correlations_teens.csv"))
write_csv(res_teens$numeric,   file.path(output_dir, "study2_table5_correlations_teens_numeric.csv"))

knitr::kable(res_adults$formatted, row.names = FALSE, caption = "Panel A. Young Adults")
```

```{r table5b, echo=FALSE}
knitr::kable(res_teens$formatted, row.names = FALSE, caption = "Panel B. Adolescents")
```

# 7. Risk subtype intercorrelations

Not a table — these open section 3.2.5. They carry the argument that the subtypes
are related but not interchangeable, which is what justifies analysing them
separately in the next section rather than collapsing them.

```{r intercorrelations}
run_corr <- function(d, x, y, cohort) {
  d2 <- d[, c(x, y)]
  d2 <- d2[complete.cases(d2), ]
  if (nrow(d2) < 3) return(data.frame(cohort = cohort, var1 = x, var2 = y, r = NA, p = NA, n = nrow(d2)))
  test <- cor.test(d2[[x]], d2[[y]], method = "pearson")
  data.frame(cohort = cohort, var1 = x, var2 = y,
             r = round(unname(test$estimate), 3),
             p = round(test$p.value, 4), n = nrow(d2))
}

subtype_pairs <- combn(risk_subtypes, 2, simplify = FALSE)

rows <- list()
for (pr in subtype_pairs) rows[[length(rows) + 1]] <- run_corr(df[is_adolescent(df), ],  pr[1], pr[2], "Adolescents")
for (pr in subtype_pairs) rows[[length(rows) + 1]] <- run_corr(df[is_young_adult(df), ], pr[1], pr[2], "Young Adults")

intercor <- bind_rows(rows)
write_csv(intercor, file.path(output_dir, "study2_risk_subtype_intercorrelations.csv"))
knitr::kable(intercor, row.names = FALSE)
```

High- and low-stakes negative risk track each other closely. Low-stakes positive
risk does not track high-stakes negative risk at all.

# 8. Shapley value regression

The predictors are correlated, so ordinary betas depend on entry order and each
predictor's apparent importance shifts with whatever else is in the model.
Shapley averages each predictor's contribution to R² across every possible
ordering, so the result sums to the model R² and does not depend on how the steps
were ordered.

Read it as share of **explained** variance, not effect size. The adolescent model
explains 20.8% of variance in MTES; risk variety accounts for about half of that
20.8%, not half of MTES.

This uses the composite rather than the three subtypes. Three heavily overlapping
risk terms would split one construct's contribution across them and make every
risk share look small. The `na.omit` is listwise by necessity — Shapley refits
over every subset of predictors and needs one constant sample to compare them.

```{r shapley}
run_shapley <- function(data, cohort_label) {
  shap_df <- data %>%
    dplyr::select(mtes_zscore, risk_variety, reward_sensitivity,
                  impulsivity, exploration, gender) %>%
    na.omit()

  shap_values <- shapleyvalue(y = shap_df$mtes_zscore,
                              x = shap_df %>% dplyr::select(-mtes_zscore) %>% as.data.frame())

  # Row 1 = raw Shapley value; row 2 = proportion of model R2, as a percent.
  shap_out <- as.data.frame(shap_values)
  shap_out <- rbind(Shapley_Value = shap_out[1, ],
                    Percent_R2    = round(as.numeric(shap_out[2, ]) * 100, 2))

  write_csv(as.data.frame(shap_out),
            file.path(output_dir, paste0("study2_shapley_", cohort_label, ".csv")))
  shap_out
}

shap_teens  <- run_shapley(df[is_adolescent(df), ],  "adolescents")
shap_adults <- run_shapley(df[is_young_adult(df), ], "adults")
knitr::kable(shap_adults, caption = "Young adults")
```

# 9. Indirect effects — Table S3

Cognitive trait → risk variety → SSMU, separately for each trait and each cohort.
Six mediations, 5000 bootstrap resamples each. Slowest section here.

Risk taking is the mediator and social media engagement is the outcome. The data
are cross-sectional, so this is statistical accounting, not causal sequence. The
paper says "accounted for" throughout and never "led to" — keep that.

Predictors are standardized *after* subsetting to cohort and *after* `na.omit`,
so everything is z-scored within the exact sample the models are fit on.
Standardizing back in section 5 across both cohorts would put adolescents and
young adults on a shared scale and change the coefficients. That is why
`risk_variety_raw` is carried forward.

```{r mediation, results='hide'}
set.seed(2025)   # without this the bootstrap CIs move on every run

run_mediation_set <- function(data, cohort_label) {
  med_df <- data %>%
    dplyr::select(mtes_zscore, risk_variety_raw, reward_sensitivity, impulsivity, exploration) %>%
    mutate(risk_variety_z       = zscore(risk_variety_raw),
           reward_sensitivity_z = zscore(reward_sensitivity),
           impulsivity_z        = zscore(impulsivity),
           exploration_z        = zscore(exploration)) %>%
    na.omit()

  # Two models per mediation: predictor -> mediator, and predictor + mediator ->
  # outcome. Written longhand rather than looped -- mediate() re-evaluates the
  # model calls internally, and building the formulas programmatically breaks it.

  med_model_imp <- lm(risk_variety_z ~ impulsivity_z, data = med_df)
  out_model_imp <- lm(mtes_zscore ~ impulsivity_z + risk_variety_z, data = med_df)
  med_imp <- mediate(med_model_imp, out_model_imp, treat = "impulsivity_z",
                     mediator = "risk_variety_z", boot = TRUE, sims = 5000)

  med_model_rs <- lm(risk_variety_z ~ reward_sensitivity_z, data = med_df)
  out_model_rs <- lm(mtes_zscore ~ reward_sensitivity_z + risk_variety_z, data = med_df)
  med_rs <- mediate(med_model_rs, out_model_rs, treat = "reward_sensitivity_z",
                    mediator = "risk_variety_z", boot = TRUE, sims = 5000)

  med_model_exp <- lm(risk_variety_z ~ exploration_z, data = med_df)
  out_model_exp <- lm(mtes_zscore ~ exploration_z + risk_variety_z, data = med_df)
  med_exp <- mediate(med_model_exp, out_model_exp, treat = "exploration_z",
                     mediator = "risk_variety_z", boot = TRUE, sims = 5000)

  # d0 = indirect (ACME), z0 = direct (ADE), n0 = proportion mediated.
  extract_results <- function(m, predictor_name) {
    data.frame(
      Predictor = predictor_name,
      Indirect  = sprintf("%.2f (%.2f, %.2f), %.3f", m$d0, m$d0.ci[1], m$d0.ci[2], m$d0.p),
      Direct    = sprintf("%.2f (%.2f, %.2f), %.2f", m$z0, m$z0.ci[1], m$z0.ci[2], m$z0.p),
      Prop      = sprintf("%.1f%%, %.2f", m$n0 * 100, m$n0.p),
      stringsAsFactors = FALSE)
  }

  out <- bind_rows(extract_results(med_imp, "impulsivity"),
                   extract_results(med_rs,  "reward_sensitivity"),
                   extract_results(med_exp, "exploration"))

  write_csv(out, file.path(output_dir, paste0("study2_mediation_", cohort_label, ".csv")))
  out
}

med_teens  <- run_mediation_set(df[is_adolescent(df), ],  "adolescents")
med_adults <- run_mediation_set(df[is_young_adult(df), ], "adults")
```

```{r mediation-show, echo=FALSE}
knitr::kable(med_adults, row.names = FALSE, caption = "Young adults")
knitr::kable(med_teens,  row.names = FALSE, caption = "Adolescents")
```

Proportion mediated exceeds 100% for reward sensitivity in both cohorts. That
happens when the direct and indirect paths have opposite signs. It is not an
error and should not be "fixed" — the paper reports it alongside a nonsignificant
p for the proportion.

# 10. Hierarchical regression — Table S4

Once you account for who someone is dispositionally, does what they do still
explain anything about their social media engagement?

1. sex + impulsivity
2. \+ reward sensitivity
3. \+ exploration
4. \+ the risk subtypes

Sex is a covariate from Step 1, not a predictor of interest. Entering the
cognitive traits before risk is the conservative ordering — it gives the
dispositional account first claim on the shared variance, so whatever risk picks
up in Step 4 is over and above traits. The ΔR² at Step 4 is the test.

Listwise deletion happens once, up front, so all four steps are fit on the same
participants. Otherwise the ΔR² comparisons would be across different samples and
`anova()` would be meaningless.

```{r hierarchical}
build_hier_table <- function(data, out_name) {
  model_vars <- c("mtes_zscore", "gender", "impulsivity", "reward_sensitivity",
                  "exploration", risk_entry_order)

  d <- data %>% dplyr::select(dplyr::all_of(model_vars)) %>% na.omit()

  base <- "mtes_zscore ~ gender + impulsivity"
  f <- c(base,
         paste(base, "+ reward_sensitivity"),
         paste(base, "+ reward_sensitivity + exploration"),
         paste(base, "+ reward_sensitivity + exploration +",
               paste(risk_entry_order, collapse = " + ")))

  models     <- lapply(f, function(x) lm(as.formula(x), data = d))
  models_std <- lapply(models, lm.beta)          # betas in the table are standardized
  anova_res  <- do.call(anova, models)           # step-to-step delta R2 tests

  format_beta <- function(beta, p) {
    stars <- if (p < .001) "***" else if (p < .01) "**" else if (p < .05) "*" else if (p < .10) "+" else ""
    sprintf("%.2f%s", beta, stars)
  }

  rows <- list()
  for (i in seq_along(models)) {
    summ <- summary(models[[i]]); summ_std <- summary(models_std[[i]])

    rows[[length(rows) + 1]] <- data.frame(Step = paste0("Step ", i), Predictor = "",
                                           Beta = "", SE = "", t = "", p = "", stringsAsFactors = FALSE)

    for (j in 2:nrow(summ$coefficients)) {       # start at 2 to skip the intercept
      rows[[length(rows) + 1]] <- data.frame(
        Step = "",
        Predictor = rownames(summ$coefficients)[j],
        Beta = format_beta(summ_std$coefficients[j, "Standardized"], summ$coefficients[j, "Pr(>|t|)"]),
        # SE and t come from the unstandardized fit; only beta is standardized.
        SE = sprintf("%.2f", summ$coefficients[j, "Std. Error"]),
        t  = sprintf("%.2f", summ$coefficients[j, "t value"]),
        p  = format_p_num(summ$coefficients[j, "Pr(>|t|)"]),
        stringsAsFactors = FALSE)
    }

    r2 <- summ$r.squared
    fstat <- summ$fstatistic
    fval <- fstat[1]; df1 <- fstat[2]; df2 <- fstat[3]
    pval <- pf(fval, df1, df2, lower.tail = FALSE)

    delta_r2 <- ""
    if (i > 1) {
      delta <- r2 - summary(models[[i - 1]])$r.squared
      delta_r2 <- paste0(", Delta R2 = ", sprintf("%.3f", delta),
                         ", p = ", format_p_num(anova_res$`Pr(>F)`[i]))
    }

    rows[[length(rows) + 1]] <- data.frame(
      Step = "",
      Predictor = paste0("R2 = ", sprintf("%.3f", r2),
                         ", F(", df1, ", ", df2, ") = ", sprintf("%.2f", fval),
                         ", p = ", format_p_num(pval), delta_r2),
      Beta = "", SE = "", t = "", p = "", stringsAsFactors = FALSE)
  }

  out <- bind_rows(rows)
  write_csv(out, file.path(output_dir, out_name))
  out
}

hier_teens  <- build_hier_table(df[is_adolescent(df), ],  "study2_table14_hierarchical_regression_adolescents.csv")
hier_adults <- build_hier_table(df[is_young_adult(df), ], "study2_table13_hierarchical_regression_adults.csv")
knitr::kable(hier_adults, row.names = FALSE, caption = "Young adults")
```

The cohorts diverge here, and it is the substantive finding: for young adults
risk does not add over traits, because impulsivity absorbs it. For adolescents
it does.

**On structuring Step 4.** All subtypes go in together. They correlate between
.25 and .55, so each beta is that subtype's unique contribution net of the
others, and none reaches significance in the young adult model even though the
block as a whole sits near threshold. Two alternatives, both fine if stated:
composite in Step 4 and subtypes in Step 5, which tests "does risk matter" before
"which kind"; or three separate Step 4 models, one subtype each, which reads each
subtype cleanly but runs three tests where the paper ran one. What is not fine is
trying each and reporting whichever makes risk significant.

# 11. Social versus parasocial — Table 6

If the link between social media and risk were just screen time, it would not
matter what kind of platform the time was spent on. It does — risk tracks
reciprocal platforms and not one-sided ones, in both cohorts.

Steiger's z is the right test because the two correlations being compared share a
variable (the risk subtype) and come from the same people. Comparing two r values
directly would ignore that dependency and overstate the difference. The test
needs a third correlation, social use with parasocial use, to know how dependent
they are.

```{r steiger}
run_steiger <- function(data) {
  rows <- list()
  for (pred in risk_subtypes) {
    r_social            <- cor(data[[pred]],    data$social_use,     use = "pairwise.complete.obs")
    r_parasocial        <- cor(data[[pred]],    data$parasocial_use, use = "pairwise.complete.obs")
    r_social_parasocial <- cor(data$social_use, data$parasocial_use, use = "pairwise.complete.obs")
    n <- nrow(na.omit(data[, c(pred, "social_use", "parasocial_use")]))

    test <- cocor.dep.groups.overlap(r.jk = r_social,             # risk x social
                                     r.jh = r_parasocial,         # risk x parasocial
                                     r.kh = r_social_parasocial,  # social x parasocial
                                     n    = n)
    steiger <- test@steiger1980

    rows[[length(rows) + 1]] <- data.frame(
      Risk_Subtype = pred,
      r_social     = round(r_social, 2),
      r_parasocial = round(r_parasocial, 2),
      # Sign flipped so a negative z means social > parasocial, matching the
      # direction printed in Table 6. cocor orders the comparison the other way.
      steiger_z    = round(-1 * steiger$statistic, 2),
      p_value      = round(steiger$p.value, 3),
      n            = n)
  }
  bind_rows(rows)
}

steiger_adults <- run_steiger(df[is_young_adult(df), ])
steiger_teens  <- run_steiger(df[is_adolescent(df), ])
write_csv(steiger_adults, file.path(output_dir, "study2_table7_steiger_adults.csv"))
write_csv(steiger_teens,  file.path(output_dir, "study2_table8_steiger_teens.csv"))

knitr::kable(steiger_adults, row.names = FALSE, caption = "Panel A. Young Adults")
```

```{r steiger-teens, echo=FALSE}
knitr::kable(steiger_teens, row.names = FALSE, caption = "Panel B. Adolescents")
```

# Running this on cohort 1

**Score it first.** One row per participant, all scales scored, outliers handled,
to the lab's standards. Then point `main_file` at it and set the right-hand side
of the column map. Ask me if any of the scoring is unclear.

**High-stakes risk.** Cohort 1 has no YRBS substance scoring. Set
`use_high_stakes <- FALSE` and every table adjusts — the row drops out of Tables
5 and 6, the term drops out of Step 4, and total risk variety becomes a
two-subtype composite weighted over 21 items instead of 28. Step 4 then tests two
risk predictors rather than three, so the ΔR² is not comparable to the published
one; report it on its own terms.

Study 1 used the BRP rather than the YRBS for this variable, so if cohort 1
carries a different high-stakes measure there is precedent in the paper for
mapping it in instead.

**Participants without a phone.** Jason and Lina have already worked out how
people who report no phone at the start of the survey are handled. Confirm with
them that it is taken care of before running anything, since the MTES composite
is z-scored across whoever is in the file.

**Cohort split.** PID still encodes age group, so `is_young_adult` and
`is_adolescent` carry over unchanged.
