---
title: "DEFT Study 2 analysis workflow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = "#>")
```

The Study 2 analyses from Zweben et al. (2026), what each one is for, and the
order they were run in, so you can do the same on cohort 1.

> Zweben, D., Zaneski, L., Chiu, M., Green, H., Brooks-Justice, J., Dachs, A.,
> Vince, M. A., Olino, T., Marshall, P., Steinberg, L., & Chein, J. M. (2026).
> Clarifying the association between digital media engagement and risk taking.
> *Technology in Society, 88*, 103491.
> [doi:10.1016/j.techsoc.2026.103491](https://doi.org/10.1016/j.techsoc.2026.103491)

Study 2 only. Two cohorts out of LITE: adolescents 11 to 17, young adults 18 to
21. Study 1 is a different sample with different instruments.

The outcome is SSMU, social media engagement, built from three MTES subscales.
The predictors are impulsivity (BIS), reward sensitivity (Zuckerman sensation
seeking), exploration (the apple-picking foraging task), and risk taking split
three ways: low-stakes positive and low-stakes negative from the PNRT, and
high-stakes negative from the YRBS substance items. Platform time is split into
social and parasocial.

Score everything and screen outliers however the lab does it. This document picks
up from a scored file, one row per participant. For reference, DEFT screened at
|z| > 3 per variable and blanked the individual cell rather than dropping the
row, so one bad foraging value did not cost that participant everything else.

The columns this expects:

| Column | What it is |
|---|---|
| `PID` | also encodes cohort: 2xxx adolescent, 3xxx+ young adult |
| `sex` | 1 female, 2 male |
| `MTES_SocialMediaUse_TimeOnApp.sum` | time on apps, summed across platforms |
| `MTES_PhoneChecking.mean` | checking subscale |
| `MTES_PublicUpdates.mean` | public posting subscale |
| `SocialUsage` / `ParasocialUsage` | time by platform type, see caveat 2 |
| `barrets.sum` | impulsivity, BIS, 8 items |
| `zuckerman.mean` | reward sensitivity, 6 items |
| `foraging` | apples per tree, raw task output |
| `positiverisk_ever_mean` / `negativerisk_ever_mean` | PNRT, see caveat 1 |
| `yrbs_var` | high-stakes negative risk. DEFT only, cohort 1 will not have it |

Two things about how we scored are not standard, and are not written down
anywhere else. They come first, because both of them affect everything after.

# Caveat 1: the PNRT

Positive risk, negative risk, and total variety all come off the PNRT, and we
score it differently than it is normally scored.

Every PNRT activity is asked twice: have you ever done this, and if so, how often
in the past six months. Variety normally comes off the second question. An
activity counts as 1 if there was any engagement in the past six months and 0 if
there was none.

We could not do that. For the first six months of LITE, before Avi and I were on
it, the frequency follow-up gave participants no way to answer zero times once
they had already said yes to ever having done the activity. For those
participants the frequency item cannot separate recent engagement from none.

So we score variety off the **ever** items, and we have kept doing that since.
The zero option may well have been fixed by now. Do not switch. Cohort 1 has to
be scored the same way for the two to be comparable.

Each score is the proportion of activities in its set that the participant has
ever done. The item lists:

```{r pnrt, eval=FALSE}
# 7 items
pnrt_neg_items <- c("pnrt_phone", "pnrt_cheat", "pnrt_skip", "pnrt_personal",
                    "pnrt_snuckout", "pnrt_sexy", "pnrt_mph")

# 14 items
pnrt_pos_items <- c("pnrt_audition", "pnrt_newclub", "pnrt_truth", "pnrt_food",
                    "pnrt_leader", "pnrt_like", "pnrt_challenge", "pnrt_newlook",
                    "pnrt_newsocial", "pnrt_secret", "pnrt_beliefs",
                    "pnrt_friend", "pnrt_sport", "pnrt_newpeople")

negativerisk_ever_mean <- rowMeans(pnrt[, pnrt_neg_items], na.rm = TRUE)
positiverisk_ever_mean <- rowMeans(pnrt[, pnrt_pos_items], na.rm = TRUE)
```

`pnrt_personal` counts as negative and `pnrt_secret` counts as positive, which is
not what the names suggest.

Total variety is the mean of positive, negative, and high-stakes negative. It
gets built in step 0 rather than at scoring, because the three pieces have to be
on the same footing first.

# Caveat 2: social versus parasocial platforms

Step 6 compares risk against each group separately, so the sorting has to be
right.

| Type | Platforms |
|---|---|
| **Social**, reciprocal, you interact with people you know | Instagram, Snapchat, Facebook |
| **Parasocial**, one-sided, you consume what strangers made | YouTube, TikTok, Pinterest, Twitter |

`SocialUsage` and `ParasocialUsage` are time summed within each group.

Cohort 1 will have platforms these lists do not cover. Sort each new one by
whether people mostly use it to interact or mostly to watch, and write down what
you decided. It goes straight into the Table 6 result.

# The analyses

Six of them, run in this order.

| Paper | Step | What it answers |
|---|---|---|
| 3.2.1 | 1. Descriptives, Table 4 | What the measures look like in each cohort |
| 3.2.2 | 2. Correlations, Table 5 | Is SSMU related to risk taking and to the traits at all |
| 3.2.3 | 3. Shapley, in text | Of the things that are related, which carries the most weight |
| 3.2.4 | 4. Indirect effects, Table S3 | Do the traits reach SSMU through risk taking |
| 3.2.5 | 5. Subtypes, then hierarchical regression, Table S4 | Which kind of risk, and does any of it survive controlling for traits |
| 3.2.6 | 6. Social vs parasocial, Table 6 | Is this about screen time, or about the kind of platform |

Two things about the order:

**Shapley runs before the regression and does not replace it.** The predictors
are correlated, so betas depend on entry order and Shapley does not. The
regression then asks whether risk adds anything once the traits have had first
claim on the shared variance.

**The subtype intercorrelations run before the subtype analyses.** They show the
three kinds of risk are related but not interchangeable, which is why the last
step splits risk into three predictors instead of using the composite.

## Cohort 1: no high-stakes risk

DEFT had low-stakes positive, low-stakes negative, and high-stakes negative.
Cohort 1 has no YRBS substance scoring, so it has the first two.

Delete the `risk_high_neg` line from the column map and the rest follows. Total
variety becomes the mean of two pieces, the high-stakes row drops out of Tables 5
and 6, and the last regression step tests two risk predictors. That ΔR² is a new
number rather than a replication of the published one, so report it that way.

Do not leave the column mapped and empty. Total variety uses `na.rm = TRUE`, so
it will average whatever is there and give you a number that looks fine.

# Setup

```{r packages, eval=FALSE}
install.packages(c("dplyr", "readr", "tidyr", "psych", "Hmisc",
                   "cocor", "mediation", "lm.beta", "ShapleyValue"))
```

```{r libraries}
suppressPackageStartupMessages({
  library(dplyr); library(readr); library(tidyr); library(psych); library(Hmisc)
  library(cocor)          # Steiger's z for dependent, overlapping correlations
  library(mediation)      # bootstrapped indirect effects
  library(lm.beta)        # standardized betas
  library(ShapleyValue)   # Shapley value regression
})

data_dir   <- "data"
output_dir <- "outputs"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
main_file  <- "study2_scored.csv"
```

Point the right-hand side at your columns. Nothing below this refers to a column
name, so this is the only thing to edit.

```{r column-map}
cols <- c(
  pid                = "PID",
  gender             = "sex",                                 # 1 female, 2 male
  mtes_time          = "MTES_SocialMediaUse_TimeOnApp.sum",
  mtes_checking      = "MTES_PhoneChecking.mean",
  mtes_updates       = "MTES_PublicUpdates.mean",
  social_use         = "SocialUsage",
  parasocial_use     = "ParasocialUsage",
  impulsivity        = "barrets.sum",
  reward_sensitivity = "zuckerman.mean",
  exploration        = "foraging",                            # raw, apples per tree
  risk_low_pos       = "positiverisk_ever_mean",
  risk_low_neg       = "negativerisk_ever_mean",
  risk_high_neg      = "yrbs_var"        # DEFT only. Delete this line for cohort 1.
)

# PID encodes age group, and that carries over to cohort 1.
is_young_adult <- function(d) d$pid > 3000
is_adolescent  <- function(d) d$pid < 3000

zscore <- function(x) (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)

# Three star levels. Study 1 additionally marked p < .10; Study 2 does not.
format_r <- function(r, p) {
  stars <- if (is.na(p)) "" else if (p < .001) "***" else if (p < .01) "**" else if (p < .05) "*" else ""
  sprintf("%.2f%s", r, stars)
}
format_p_num <- function(p) if (p < .001) "< .001" else sprintf("%.3f", p)
```

# Step 0: Variables the analysis builds

Everything else arrives scored. These three are analysis decisions, so they
happen here.

**SSMU composite.** The outcome for the whole paper. The three MTES subscales sit
on different scales, since time is a sum across platforms and the other two are
item means, so each gets z-scored before averaging. Otherwise time dominates.

It is z-scored within whoever is in the file, so it is not on the same absolute
scale across datasets. Compare patterns of association, not means. It also has to
be built before anything is joined on, since a join that changes the row count
moves the mean and SD with it.

**Exploration.** Apples per tree. Staying on a patch means exploiting, so higher
raw values mean *less* exploration. Sign-flipped so the variable reads in the
direction of its name. Zeros become NA, since someone who harvested nothing did
not do the task.

**Total risk variety.** The mean of the risk subtypes, then z-scored. The
unstandardized version is kept because step 4 re-standardizes within its own
subsample.

```{r step0}
raw <- readr::read_csv(file.path(data_dir, main_file), show_col_types = FALSE)

missing_cols <- setdiff(unname(cols), names(raw))
if (length(missing_cols)) stop("Not in ", main_file, ": ", paste(missing_cols, collapse = ", "))

df <- raw %>%
  dplyr::select(dplyr::all_of(unname(cols))) %>%
  setNames(names(cols)) %>%
  mutate(across(everything(), as.numeric)) %>%
  mutate(mtes_zscore = rowMeans(cbind(zscore(mtes_time), zscore(mtes_checking),
                                      zscore(mtes_updates)), na.rm = TRUE),
         exploration = ifelse(exploration == 0, NA, exploration),
         exploration = -1 * exploration)

# Whichever risk subtypes you mapped, in the order the tables list them. Drop a
# line from the map above and everything downstream follows.
risk_subtypes <- intersect(c("risk_high_neg", "risk_low_neg", "risk_low_pos"), names(df))

# Each subtype is a proportion over a different number of items, so the Table 4
# percentage is weighted by item count rather than a plain mean of the three.
risk_item_counts <- c(risk_high_neg = 7, risk_low_neg = 7, risk_low_pos = 14)

# The tables list subtypes high-stakes first; the regression entered them the
# other way round. Order does not change the fit, only the printed row order.
risk_entry_order <- rev(risk_subtypes)

df <- df %>%
  mutate(risk_variety_raw = rowMeans(dplyr::select(., dplyr::all_of(risk_subtypes)), na.rm = TRUE),
         risk_variety     = zscore(risk_variety_raw))

nrow(df)
```

# Step 1: Descriptives (Table 4)

On original scales, so this comes before anything gets standardized. The
published table reports impulsivity as a BIS mean on the 1 to 4 scale while the
models use the sum. This reports whatever you supplied, so say which in your
table note.

```{r step1}
w <- risk_item_counts[risk_subtypes]
df$total_risk_pct <- as.numeric(as.matrix(df[, risk_subtypes]) %*% w / sum(w) * 100)

summarise_var <- function(x) {
  data.frame(Mean = round(mean(x, na.rm = TRUE), 2), SD = round(sd(x, na.rm = TRUE), 2),
             Min = round(min(x, na.rm = TRUE), 2), Max = round(max(x, na.rm = TRUE), 2))
}

# Row order matches Table 4 top to bottom.
desc_vars <- c("mtes_time", "mtes_updates", "mtes_checking", "social_use", "parasocial_use",
               "impulsivity", "reward_sensitivity", "total_risk_pct", risk_subtypes, "exploration")

rows <- list()
for (cohort_name in c("Young Adults", "Adolescents")) {
  dfc <- if (cohort_name == "Young Adults") df[is_young_adult(df), ] else df[is_adolescent(df), ]
  for (v in desc_vars) {
    values <- dfc[[v]]
    if (v %in% risk_subtypes) values <- values * 100   # printed as percentages
    rows[[length(rows) + 1]] <- data.frame(Cohort = cohort_name, Variable = v,
                                           summarise_var(values), stringsAsFactors = FALSE)
  }
}

table4 <- bind_rows(rows)
write_csv(table4, file.path(output_dir, "study2_table4_descriptives.csv"))
knitr::kable(table4[table4$Cohort == "Young Adults", ], row.names = FALSE)
```

Exploration shows here on the raw apples-per-tree scale, before the sign flip.

The two panel headings in the published Table 4 are transposed. Its "Young
Adults" panel holds the adolescent numbers. Age confirms the split is right
(PID > 3000 is 18 to 21, PID < 3000 is 11 to 17), and Table 5 is labelled
correctly, so it is only Table 4.

# Step 2: Correlations (Table 5)

One published table, two panels. Pairwise complete rather than listwise, so each
cell uses everyone who has both variables. That is why the table carries stars
instead of a single N.

```{r step2}
main_vars <- c("mtes_zscore", "impulsivity", "reward_sensitivity", "exploration")
row_vars  <- c(main_vars, "risk_variety", risk_subtypes)

get_r_p <- function(d, x, y) {
  d2 <- d[, c(x, y)]; d2 <- d2[complete.cases(d2), ]
  if (nrow(d2) < 3) return(list(r = NA_real_, p = NA_real_))
  t <- cor.test(d2[[x]], d2[[y]], method = "pearson")
  list(r = unname(t$estimate), p = t$p.value)
}

build_corr_table <- function(d) {
  r_mat <- matrix(NA_real_, length(row_vars), length(main_vars),
                  dimnames = list(row_vars, main_vars))
  p_mat <- r_mat
  for (rv in row_vars) for (cv in main_vars) {
    if (rv == cv) next                       # diagonal stays blank, as in the paper
    res <- get_r_p(d, rv, cv); r_mat[rv, cv] <- res$r; p_mat[rv, cv] <- res$p
  }
  out <- matrix("", nrow(r_mat), ncol(r_mat), dimnames = dimnames(r_mat))
  for (i in seq_len(nrow(r_mat))) for (j in seq_len(ncol(r_mat)))
    if (!is.na(r_mat[i, j])) out[i, j] <- format_r(r_mat[i, j], p_mat[i, j])
  as.data.frame(cbind(Variable = rownames(out), out))
}

t5_adults <- build_corr_table(df[is_young_adult(df), ])
t5_teens  <- build_corr_table(df[is_adolescent(df), ])
write_csv(t5_adults, file.path(output_dir, "study2_table5_correlations_adults.csv"))
write_csv(t5_teens,  file.path(output_dir, "study2_table5_correlations_teens.csv"))
knitr::kable(t5_adults, row.names = FALSE, caption = "Panel A. Young Adults")
```

```{r step2b, echo=FALSE}
knitr::kable(t5_teens, row.names = FALSE, caption = "Panel B. Adolescents")
```

**Cohort 1:** the high-stakes row disappears.

# Step 3: Shapley

The predictors are correlated, so ordinary betas shift depending on what else is
in the model and in what order. Shapley averages each predictor's contribution to
R² over every possible ordering, so it sums to the model R² and does not depend
on entry order.

Read it as share of *explained* variance, not effect size. The adolescent model
explains 20.8% of the variance in MTES, and risk variety accounts for about half
of that 20.8%, not half of MTES.

It uses total variety rather than the three subtypes. Three overlapping risk
terms would split one construct's contribution across them and make every risk
share look small. The `na.omit` has to be listwise, since Shapley refits over
every subset of predictors and needs one constant sample.

First the subtype intercorrelations. They are reported in the text, and they are
why the subtypes get treated separately later.

```{r step3-intercor}
run_corr <- function(d, x, y, cohort) {
  d2 <- d[, c(x, y)]; d2 <- d2[complete.cases(d2), ]
  if (nrow(d2) < 3) return(data.frame(cohort, var1 = x, var2 = y, r = NA, p = NA, n = nrow(d2)))
  t <- cor.test(d2[[x]], d2[[y]], method = "pearson")
  data.frame(cohort, var1 = x, var2 = y, r = round(unname(t$estimate), 3),
             p = round(t$p.value, 4), n = nrow(d2))
}

rows <- list()
for (pr in combn(risk_subtypes, 2, simplify = FALSE)) {
  rows[[length(rows) + 1]] <- run_corr(df[is_adolescent(df), ],  pr[1], pr[2], "Adolescents")
  rows[[length(rows) + 1]] <- run_corr(df[is_young_adult(df), ], pr[1], pr[2], "Young Adults")
}
intercor <- bind_rows(rows)
write_csv(intercor, file.path(output_dir, "study2_risk_subtype_intercorrelations.csv"))
knitr::kable(intercor, row.names = FALSE)
```

```{r step3-shapley}
run_shapley <- function(data, label) {
  sd_ <- data %>%
    dplyr::select(mtes_zscore, risk_variety, reward_sensitivity,
                  impulsivity, exploration, gender) %>%
    na.omit()
  sv <- shapleyvalue(y = sd_$mtes_zscore,
                     x = sd_ %>% dplyr::select(-mtes_zscore) %>% as.data.frame())
  out <- as.data.frame(sv)
  # Row 1 = raw Shapley value; row 2 = share of model R2, as a percent.
  out <- rbind(Shapley_Value = out[1, ], Percent_R2 = round(as.numeric(out[2, ]) * 100, 2))
  write_csv(as.data.frame(out), file.path(output_dir, paste0("study2_shapley_", label, ".csv")))
  out
}

shap_teens  <- run_shapley(df[is_adolescent(df), ],  "adolescents")
shap_adults <- run_shapley(df[is_young_adult(df), ], "adults")
knitr::kable(shap_adults, caption = "Young adults")
```

**Cohort 1:** total variety is now two subtypes, so its share is not comparable
to the published one.

# Step 4: Indirect effects (Table S3)

Trait to risk variety to SSMU, separately for each trait and each cohort. Six
mediations at 5000 bootstrap resamples each. This one takes a few minutes.

Risk taking is the mediator and SSMU is the outcome. The data are
cross-sectional, so this is statistical accounting rather than causal order. The
paper says "accounted for" throughout and never "led to". Keep that.

Predictors get standardized after subsetting to cohort and after `na.omit`, so
everything is z-scored within the exact sample the models are fit on.
Standardizing earlier, across both cohorts, would put adolescents and young
adults on a shared scale and change the coefficients.

```{r step4, results='hide'}
set.seed(2025)   # bootstrap CIs move on every run without this

run_mediation_set <- function(data, label) {
  m <- data %>%
    dplyr::select(mtes_zscore, risk_variety_raw, reward_sensitivity, impulsivity, exploration) %>%
    mutate(risk_variety_z = zscore(risk_variety_raw), reward_sensitivity_z = zscore(reward_sensitivity),
           impulsivity_z = zscore(impulsivity), exploration_z = zscore(exploration)) %>%
    na.omit()

  # Written out longhand rather than looped. mediate() re-evaluates the model
  # calls internally, and building the formulas programmatically breaks it.
  mi <- mediate(lm(risk_variety_z ~ impulsivity_z, data = m),
                lm(mtes_zscore ~ impulsivity_z + risk_variety_z, data = m),
                treat = "impulsivity_z", mediator = "risk_variety_z", boot = TRUE, sims = 5000)
  mr <- mediate(lm(risk_variety_z ~ reward_sensitivity_z, data = m),
                lm(mtes_zscore ~ reward_sensitivity_z + risk_variety_z, data = m),
                treat = "reward_sensitivity_z", mediator = "risk_variety_z", boot = TRUE, sims = 5000)
  me <- mediate(lm(risk_variety_z ~ exploration_z, data = m),
                lm(mtes_zscore ~ exploration_z + risk_variety_z, data = m),
                treat = "exploration_z", mediator = "risk_variety_z", boot = TRUE, sims = 5000)

  # d0 = indirect (ACME), z0 = direct (ADE), n0 = proportion mediated.
  ex <- function(x, nm) data.frame(
    Predictor = nm,
    Indirect  = sprintf("%.2f (%.2f, %.2f), %.3f", x$d0, x$d0.ci[1], x$d0.ci[2], x$d0.p),
    Direct    = sprintf("%.2f (%.2f, %.2f), %.2f", x$z0, x$z0.ci[1], x$z0.ci[2], x$z0.p),
    Prop      = sprintf("%.1f%%, %.2f", x$n0 * 100, x$n0.p), stringsAsFactors = FALSE)

  out <- bind_rows(ex(mi, "impulsivity"), ex(mr, "reward_sensitivity"), ex(me, "exploration"))
  write_csv(out, file.path(output_dir, paste0("study2_mediation_", label, ".csv")))
  out
}

med_teens  <- run_mediation_set(df[is_adolescent(df), ],  "adolescents")
med_adults <- run_mediation_set(df[is_young_adult(df), ], "adults")
```

```{r step4-show, echo=FALSE}
knitr::kable(med_adults, row.names = FALSE, caption = "Young adults")
knitr::kable(med_teens,  row.names = FALSE, caption = "Adolescents")
```

Proportion mediated goes over 100% for reward sensitivity in both cohorts. That
happens when the direct and indirect paths have opposite signs. It is not an
error and should not be "fixed". The paper reports it with a nonsignificant p for
the proportion.

# Step 5: Hierarchical regression (Table S4)

Once you account for who someone is dispositionally, does what they do still
explain anything about their media engagement?

1. sex + impulsivity
2. \+ reward sensitivity
3. \+ exploration
4. \+ the risk subtypes

Sex is a covariate from the start, not a predictor of interest. Traits go in
before risk on purpose. That gives the dispositional account first claim on the
shared variance, so anything risk picks up in the last step is over and above
them. That ΔR² is the test.

Listwise deletion happens once, up front, so all four steps are fit on the same
people. Otherwise the ΔR² comparisons run across different samples and `anova()`
means nothing.

```{r step5}
build_hier_table <- function(data, out_name) {
  model_vars <- c("mtes_zscore", "gender", "impulsivity", "reward_sensitivity",
                  "exploration", risk_entry_order)
  d <- data %>% dplyr::select(dplyr::all_of(model_vars)) %>% na.omit()

  base <- "mtes_zscore ~ gender + impulsivity"
  f <- c(base, paste(base, "+ reward_sensitivity"),
         paste(base, "+ reward_sensitivity + exploration"),
         paste(base, "+ reward_sensitivity + exploration +",
               paste(risk_entry_order, collapse = " + ")))

  models     <- lapply(f, function(x) lm(as.formula(x), data = d))
  models_std <- lapply(models, lm.beta)          # table betas are standardized
  anova_res  <- do.call(anova, models)

  fb <- function(b, p) {
    s <- if (p < .001) "***" else if (p < .01) "**" else if (p < .05) "*" else if (p < .10) "+" else ""
    sprintf("%.2f%s", b, s)
  }

  rows <- list()
  for (i in seq_along(models)) {
    su <- summary(models[[i]]); ss <- summary(models_std[[i]])
    rows[[length(rows) + 1]] <- data.frame(Step = paste0("Step ", i), Predictor = "",
                                           Beta = "", SE = "", t = "", p = "", stringsAsFactors = FALSE)
    for (j in 2:nrow(su$coefficients)) {         # start at 2 to skip the intercept
      rows[[length(rows) + 1]] <- data.frame(
        Step = "", Predictor = rownames(su$coefficients)[j],
        Beta = fb(ss$coefficients[j, "Standardized"], su$coefficients[j, "Pr(>|t|)"]),
        # SE and t come from the unstandardized fit; only beta is standardized.
        SE = sprintf("%.2f", su$coefficients[j, "Std. Error"]),
        t  = sprintf("%.2f", su$coefficients[j, "t value"]),
        p  = format_p_num(su$coefficients[j, "Pr(>|t|)"]), stringsAsFactors = FALSE)
    }
    r2 <- su$r.squared; fs <- su$fstatistic
    pv <- pf(fs[1], fs[2], fs[3], lower.tail = FALSE)
    dr <- if (i > 1) paste0(", Delta R2 = ", sprintf("%.3f", r2 - summary(models[[i-1]])$r.squared),
                            ", p = ", format_p_num(anova_res$`Pr(>F)`[i])) else ""
    rows[[length(rows) + 1]] <- data.frame(
      Step = "", Predictor = paste0("R2 = ", sprintf("%.3f", r2), ", F(", fs[2], ", ", fs[3], ") = ",
                                    sprintf("%.2f", fs[1]), ", p = ", format_p_num(pv), dr),
      Beta = "", SE = "", t = "", p = "", stringsAsFactors = FALSE)
  }
  out <- bind_rows(rows); write_csv(out, file.path(output_dir, out_name)); out
}

hier_teens  <- build_hier_table(df[is_adolescent(df), ],  "study2_tableS4_adolescents.csv")
hier_adults <- build_hier_table(df[is_young_adult(df), ], "study2_tableS4_adults.csv")
knitr::kable(hier_adults, row.names = FALSE, caption = "Young adults")
```

The cohorts differ here, and that is the main finding. For young adults risk
does not add over traits, because impulsivity absorbs it. For adolescents it
does.

All the subtypes go into the last step together. They correlate between .25 and
.55, so each beta is that subtype's unique contribution net of the others, and
none of them reaches significance in the young adult model even though the block
as a whole sits near threshold. Two other defensible ways to do it: total variety
in step 4 with the subtypes in a step 5, which asks whether risk matters before
asking which kind; or three separate step 4 models, one subtype each, which reads
each subtype cleanly but runs three tests where the paper ran one. Either is fine
as long as you say which you did. Do not run all of them and report whichever one
makes risk significant.

**Cohort 1:** two risk predictors in the last step instead of three, so that ΔR²
is a new result rather than a replication.

# Step 6: Social versus parasocial (Table 6)

If the effect were just screen time, platform type would not matter. It does.
Risk tracks reciprocal platforms and not one-sided ones, in both cohorts.

Steiger's z is the right test because the two correlations being compared share a
variable, the risk subtype, and come from the same people. Comparing two r values
directly ignores that and overstates the difference. The test needs a third
correlation, social with parasocial, to know how dependent they are.

```{r step6}
run_steiger <- function(data) {
  rows <- list()
  for (pred in risk_subtypes) {
    r_s  <- cor(data[[pred]], data$social_use,     use = "pairwise.complete.obs")
    r_p  <- cor(data[[pred]], data$parasocial_use, use = "pairwise.complete.obs")
    r_sp <- cor(data$social_use, data$parasocial_use, use = "pairwise.complete.obs")
    n <- nrow(na.omit(data[, c(pred, "social_use", "parasocial_use")]))
    st <- cocor.dep.groups.overlap(r.jk = r_s, r.jh = r_p, r.kh = r_sp, n = n)@steiger1980
    rows[[length(rows) + 1]] <- data.frame(
      Risk_Subtype = pred, r_social = round(r_s, 2), r_parasocial = round(r_p, 2),
      # Sign flipped so a negative z means social > parasocial, matching Table 6.
      # cocor orders the comparison the other way.
      steiger_z = round(-1 * st$statistic, 2), p_value = round(st$p.value, 3), n = n)
  }
  bind_rows(rows)
}

st_adults <- run_steiger(df[is_young_adult(df), ])
st_teens  <- run_steiger(df[is_adolescent(df), ])
write_csv(st_adults, file.path(output_dir, "study2_table6_steiger_adults.csv"))
write_csv(st_teens,  file.path(output_dir, "study2_table6_steiger_teens.csv"))
knitr::kable(st_adults, row.names = FALSE, caption = "Panel A. Young Adults")
```

```{r step6b, echo=FALSE}
knitr::kable(st_teens, row.names = FALSE, caption = "Panel B. Adolescents")
```

**Cohort 1:** the high-stakes row drops, and this step depends on the platform
sorting in caveat 2.

# Before you start

- Score the PNRT off the ever items, using the lists in caveat 1.
- Sort cohort 1's platforms into social and parasocial, and write down the calls.
- Jason and Lina have already handled how people who report no phone at the start
  of the survey are treated. Check that it is sorted in the cohort 1 export
  before you run anything, since the SSMU composite is z-scored across whoever is
  in the file.
- Delete the `risk_high_neg` line from the column map.

Then run it top to bottom. Ask me when something does not look right.
